\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}

\bibliographystyle{plainurl}% the mandatory bibstyle

\input{mycommands-CSL.tex}
\input{mypackages}
\usepackage{doc}
\usepackage{tikz}
\usepackage{todonotes}
\newcommand{\notetwo}[3]{\todo[size=\tiny,color=#1]{\texttt{\color{white} #2: #3}}}
\newcommand\co[1]{\notetwo{orange}{Carlos}{#1}}
\newcommand\ep[1]{\notetwo{blue}{Elaine}{#1}}

\usetikzlibrary{positioning, quotes}

\title{Playing with modalities}

\titlerunning{Playing with modalities}


\author{Elaine Pimentel\footnote{Corresponding author.}}{Computer Science Department UCL, UK \and \url{https://sites.google.com/site/elainepimentel/} }{e.pimentel@ucl.ac.uk}{https://orcid.org/0000-0002-7113-0801}{Pimentel has received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Sk\l odowska-Curie grant agreement Number 101007627 and by the Leverhulme Project ECUMENICAL.}

\author{Carlos Olarte}{LIPN, CNRS UMR 7030, Universit\'{e} Sorbonne Paris Nord, France \and \url{https://sites.google.com/site/carlosolarte/} }{olarte@lipn.univ-paris13.fr}{https://orcid.org/0000-0002-7264-7773}{The work of Olarte has been partially supported by the SGR project PROMUEVA (BPIN
2021000100160) under the supervision of Minciencias (Ministerio de Ciencia Tecnolog\'ia e Innovaci\'on, Colombia). Olarte acknowledges also support from the NATO
Science for Peace
and Security Programme through grant number G6133 (project SymSafe). }

\author{Timo Lang}{Computer Science Department UCL, UK \and \url{https://www.timolang.com/}}{timo.lang@ucl.ac.uk}{0000-0002-8257-968X}{}

\author{Robert Freiman}{TU-Wien, Austria}{robert@logic.at}{0000-0001-8251-4272}{}

\author{Christian G. Ferm\"{u}ller}{TU-Wien, Austria \and \url{https://www.logic.at/staff/chrisf/home.html}}{chrisf@logic.at}{0000-0003-2932-5477}{}

\authorrunning{E. Pimentel, C. Olarte, T. Lang, R. Freiman and C. Fehrm\"{u}ller} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{E. Pimentel, C. Olarte, T. Lang, R. Freiman and C. Fehrm\"{u}ller} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003752.10003790.10003801</concept_id>
       <concept_desc>Theory of computation~Linear logic</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003790.10003793</concept_id>
       <concept_desc>Theory of computation~Modal and temporal logics</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003790.10003792</concept_id>
       <concept_desc>Theory of computation~Proof theory</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Linear logic}
\ccsdesc[500]{Theory of computation~Modal and temporal logics}
\ccsdesc[500]{Theory of computation~Proof theory}
%
%\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm 

\keywords{Linear logic, modal logic, proof theory, game semantics} %TODO mandatory; please add comma-separated list of keywords

\category{Invited Talk} 

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%\acknowledgements{I want to thank \dots}

\EventEditors{J\"{o}rg Endrullis and Sylvain Schmitz}
\EventNoEds{2}
\EventLongTitle{33rd EACSL Annual Conference on Computer Science Logic (CSL 2025)}
\EventShortTitle{CSL 2025}
\EventAcronym{CSL}
\EventYear{2025}
\EventDate{February 10--14, 2025}
\EventLocation{Amsterdam, Netherlands}
\EventLogo{}
\SeriesVolume{326}
\ArticleNo{4}

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
In this work, we will explore modalities through dialogical game lenses. Games provide a powerful tool for bridging the gap between intended and formal semantics, often offering a more conceptually natural approach to logic than traditional model-theoretic semantics.

We begin by exploring substructural calculi  from a game semantic perspective, driven by intuitions about resource-consciousness and, more specifically, cost-sensitive reasoning. The game comes into full swing as we introduce cost labels to assumptions and a corresponding budget. Different proofs of the same end-sequent are interpreted as strategies for a player to defend a claim, which vary in cost. This leads to a labelled calculus, which can be viewed as a fragment of subexponential linear logic. 
%
We conclude this first part with a discussion of cut-admissibility for the proposed system. 

In the second part, we show that our games offer an interesting insight also into modal logics. More precisely, we will focus on the modal logic \PNL, characterized by Kripke frames with two types of disjoint and symmetric reachability relations. This framework is motivated by the study of group polarization, where the opinions or beliefs of individuals within a group become more extreme or polarized after interaction. Our approach to reasoning about group polarization is based on \PNL\ and highlights a different aspect of formal reasoning about the corresponding models -- using games and proof systems.
%
We conclude by outlining potential directions for future research.

\end{abstract}

\section{Introduction}\label{sec:intro}
\input{intro}

\section{A game model for costs}\label{sec:sell}
\input{sell}
\subsection{About cut-admissibility}\label{subsec:cut}
\input{cut}
\subsection{Discussion -- part I}\label{subsec:conc1}
\input{conc1}
\section{A game model for polarisation}
\input{modal}
\subsection{Discussion -- part II}\label{subsec:conc2}
\input{conc2}
\bibliography{references}
\end{document}

\end{document}
